Nodal Discontinuous Galerkin Method for the Euler Equations of Fluid Dynamics

### 2017-01-09 Documents/Euler/dgEulerCurvilinear.tex
Updates to discretization document for Euler

\newcommand{\ee}[1]{{\color{red} #1}}
\newcommand{\dx}{\Delta x}
\newcommand{\pbK}{\partial\mathbf{K}}
\newcommand{\UDG}{\bU_{\mbox{\tiny DG}}}
\newcommand{\sumx}{\sum_{i=1}^{d}}

@@

The model given by Eqs.~\eqref{eq:massConservation} - \eqref{eq:electronConservation} is non-relativistic (i.e., it does not include effects due to strong gravitational fields or relativistic fluid velocities), but is expressed in terms of curvilinear coordinates through the spatial metric tensor $\gamma_{ij}$, which gives squared infinitesimal line-element
\begin{equation}
  ds_{\vect{x}}^{2}=\gamma_{ij}\,dx^{i}\,dx^{j},
\end{equation}
where $\gamma_{ij}$ are the spatial components of the coordinate basis metric tensor.  
We will make two simplifying assumptions: the metric tensor is (1) diagonal and (2) time-independent.  
Specifically we assume the following diagonal form
\begin{equation}
  \gamma_{ij}
  =\mbox{diag}\big[\,\gamma_{11},\,\gamma_{22}(x^{1}),\,\gamma_{33}(x^{1},x^{2})\,\big]
  =\mbox{diag}\big[\,1,\,a(x^{1})^{2},\,b(x^{1})^{2}c(x^{2})^{2}\,\big], 
  \label{eq:threeMetric}
\end{equation}
which is sufficiently general to accommodate Cartesian, cylindrical, and spherical polar coordinates (see Table~\ref{tab:metricFunctions} and Appendix~\ref{app:CurvilinearEuler}).  
Moreover, $\gamma$ is the determinant of the metric tensor, and $\sqrt{\gamma}=abc$.  
The inverse of the spatial metric is denoted $\gamma^{ij}$, so that $\gamma^{ik}\gamma_{kj}=\delta^{i}_{~j}$.  
We can use the spatial metric to raise and lower indices on spatial vectors and tensors; e.g., $S_{j}=\gamma_{jk}\,S^{k}$.

\begin{table}
  \begin{center}
  \caption{Metric functions for Cartesian, Cylindrical, and Spherical coordinate systems.\label{tab:metricFunctions}}
  \begin{tabular}{cccccccccccccc}
    \midrule
    Coordinates & $x^{1}$ & $x^{2}$ & $x^{3}$ & $\bar{w}^{1}$ & $\bar{w}^{2}$ & $\bar{w}^{3}$ & $a$ & $b$ & $c$ & $\sqrt{\gamma}$
    & $\pderiv{a}{x^{1}}$ & $\pderiv{b}{x^{1}}$ & $\pderiv{c}{x^{2}}$ \\
    \midrule
    \midrule
    Cartesian & $x$ & $y$ & $z$ & $w_{x}$ & $w_{y}$ & $w_{z}$ & 1 & 1 & 1 & 1 & 0 & 0 & 0 \\
    Cylindrical & $R$ & $z$ & $\phi$ & $w_{R}$ & $w_{z}$ & $w_{\phi}$ & 1 & $R$ & 1 & $R$ & 0 & 1 & 0 \\
    Spherical & $r$ & $\theta$ & $\phi$ & $w_{r}$ & $w_{\theta}$ & $w_{\phi}$ & $r$ & $r$ & $\sin\theta$ & $r^{2}\sin\theta$ & 1 & 1 & $\cos\theta$ \\
    \midrule
    \midrule
  \end{tabular}
  \end{center}
\end{table}

@@

We define the geometry sources (fictitious plus gravitational sources) in Eq.~\eqref{eq:momentumConservation} by
\begin{equation}
  G_{j}
  =\f{1}{2}\,P^{ik}\,\pderiv{\gamma_{ik}}{x^{j}}-\rho\,\pderiv{\Phi}{x^{j}}.  
\end{equation}
Specifically, using the time-independent metric tensor in Eq.~\eqref{eq:threeMetric}, the components are
\begin{equation}
  G_{1}
  =\f{1}{2}\,\Big(\,P^{22}\,\pderiv{\gamma_{22}}{x^{1}}+P^{33}\,\pderiv{\gamma_{33}}{x^{1}}\,\Big)-\rho\,\pderiv{\Phi}{x^{1}}, \quad
  G_{2}
  =\f{1}{2}\,P^{33}\,\pderiv{\gamma_{33}}{x^{2}}-\rho\,\pderiv{\Phi}{x^{2}}, \quad\text{and}\quad
  G_{3}
  =-\rho\,\pderiv{\Phi}{x^{3}}.
\end{equation}
The gravitational energy source term in Eq.~\eqref{eq:energyConservation} is denoted with $G_{E}=-S^{i}\,\pd{\Phi}{x^{i}}$.

Then, by defining the vector of evolved variables $\bU=\big(\,\rho,\,S_{1},\,S_{2},\,S_{3},\,E,\,n_{e}\,\big)^{T}$, the vector of geometry sources $\bG(\bU)=\big(\,0,\,G_{1},\,G_{2},\,G_{3},\,G_{E},\,0\,\big)^{T}$, and the flux vector components
\begin{equation}
  \vect{F}^{i}(\bU)
  =\big(\,\rho\,v^{i},\,P_{~1}^{i},\,P_{~2}^{i},\,P_{~3}^{i},\,(E+p)\,\bar{v}^{i},\,n_{e}\,\bar{v}^{i}\,\big)^{T},
\end{equation}
we write the system of equations in the usual compact form
\begin{equation}
  \pd{\bU}{t}
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,\vect{F}^{i}(\bU)\,\Big)
  =\bG(\bU).
  \label{eq:extendedEulerCompact}
\end{equation}
Eq.~\eqref{eq:extendedEulerCompact} forms the basis for developing DG methods for the Euler equations.

@@

@@

We divide the computational domain $D$ into a disjoint union $\mathcal{T}$ of open elements $\bK$, so that $D = \cup_{\bK \in \mathcal{T}}\bK$.  
We require that each element is a box in the logical coordinates; i.e.,
\begin{equation}
  \bK=\{\vect{x} : x^{i} \in K^{i} := (\xL^{i},\xH^{i})\}, 
\end{equation}
with the surface elements denoted $\pbK^{i}=\otimes_{j\ne i}K^{j}$.  
We use $V_{\bK}$ to denote the proper volume of the element
\begin{equation}
  V_{\bK} = \int_{\bK}dV, \quad\text{where}\quad dV = \sqrt{\gamma}\,\prod_{i=1}^{d}dx^{i}.  
\end{equation}
We also define $\bx=\{\tilde{\bx}^{i},x^{i}\}$ and $\dx^{i}=\xH^{i}-\xL^{i}$.  
Here we let the approximation space for the DG method, $\mathbb{V}^{k}$, to be constructed from tensor products of one-dimensional polynomials of maximal degree $k$.  
Note that functions in $\mathbb{V}^{k}$ can be discontinuous across element interfaces.

The semi-discrete DG problem is to find $\UDG\in\mathbb{V}^{k}$ (which approximates $\bU$) such that
\begin{align}
  &\partial_{t}\int_{\bK}\UDG\,v\,dV
  +\sumx\int_{\pbK^{i}}\big(\,v\,\widehat{\vect{F}}^{i}(\UDG)\,dA^{i}\big|_{\xH^{i}}-v\,\widehat{\vect{F}}^{i}(\UDG)\,dA^{i}\big|_{\xL^{i}}\,\big) \nonumber \\
  &\hspace{24pt}
  -\sumx\int_{\bK}\vect{F}^{i}(\UDG)\,\nabla_{i}v\,dV
  =\int_{\bK}\vect{S}(\UDG)\,v\,dV
  \label{eq:semidiscreteDG}
\end{align}
for all $v\in\mathbb{V}^{k}$ and all $\bK\in\mathcal{T}$.  
Here, $dV=abc\,dx^{1}\,dx^{2}\,dx^{3}$, the area elements are given by
\begin{equation}
  dA^{1}=abc\,dx^{2}\,dx^{3}=\tau^{1}\,dx^{2}\,dx^{3}, \quad
  dA^{2}=bc\,dx^{1}\,dx^{3}=\tau^{2}\,dx^{1}\,dx^{3}, \quad
  dA^{3}=a\,dx^{1}\,dx^{2}=\tau^{3}\,dx^{1}\,dx^{2}, 
\end{equation}
and the gradient operator is $\nabla_{i}=e^{j}_{~i}\pderiv{}{x^{j}}$.

\ee{Write more about Riemann solvers here.}
In \eqref{eq:semidiscreteDG}, $\vect{\widehat{F}}^{i}(\UDG)$ is a numerical flux approximating the flux on the $i$th surface of $\bK$.  
We also define the numerical flux function $\vect{\mathcal{F}}^{i}$, which for any $v\in\mathbb{V}^{k}$ is written as
\begin{equation}
  \vect{\widehat{F}}^{i}(v)=\vect{\mathcal{F}}^{i}(v(x^{i,-},\tilde{\bx}^{i}),v(x^{i,+},\tilde{\bx}^{i})),
\end{equation}
where superscripts $-/+$, e.g., in the arguments of $v(x^{i,-/+},\tilde{\bx}^{i})$, indicate that the function is evaluated to the immediate left/right of $x^{i}$.

The approximation space $\mathbb{V}^{k}$ contains the constant functions, and the choice $v=1$ in \eqref{eq:semidiscreteDG} gives
\begin{equation}
  \partial_{t}\bU_{\bK}
  +\f{1}{V_{\bK}}\sumx\int_{\pbK^{i}}\big(\,\widehat{\vect{F}}^{i}(\UDG)\,dA^{i}\big|_{\xH^{i}}-\widehat{\vect{F}}^{i}(\UDG)\,dA^{i}\big|_{\xL^{i}}\,\big)
  =\bS_{\bK},
\end{equation}
where we have defined the volume averages
\begin{equation}
  \bU_{\bK}=\f{1}{V_{\bK}}\int_{\bK}\UDG\,dV \quad\text{and}\quad
  \bS_{\bK}=\f{1}{V_{\bK}}\int_{\bK}\vect{S}(\UDG)\,dV.  
\end{equation}
Here we use explicit, SSP-RK time integrators, which are convex combinations of forward Euler time steps \citep{gottlieb_etal_2001}.  
Thus, without loss of generality, we consider here only a forward Euler time step for $\bU_{\bK}$; i.e., 
\begin{equation}
  \bU_{\bK}^{n+1}=\bU_{\bK}^{n}
  =-\f{\dt}{V_{\bK}}\sumx\int_{\pbK^{i}}\big(\,\widehat{\vect{F}}^{i}(\UDG^{n})\,dA^{i}\big|_{\xH^{i}}-\widehat{\vect{F}}^{i}(\UDG^{n})\,dA^{i}\big|_{\xL^{i}}\,\big)
  +\vect{S}_{\bK}^{n},
\end{equation}
where $\dt=t^{n+1}-t^{n}$ is the time step.

@@

In each element $\bK$, we use a nodal representation in the conserved variables $\bU$; i.e.,
\begin{equation}
  \bU(\bx,t)\approx
  \bU_{\DG}(\bx,t)=\sum_{\bi=\vect{1}}^{\bN}\bU_{\bi}(t)\,\ell_{\bi}(\bx),
  \label{eq:conservedNodalExpansion}
\end{equation}
where $\ell_{\bi}(\bx)\in\bbQ^{k}$ are basis functions.  
(Specifically, we use Lagrange polynomials $\ell_{i}(x)$ to construct the multidimensional representation.)  
Moreover $\bi=\{i_{1},\ldots,i_{d}\}$ is a multi-index, and $\ell_{\bi}(\bx)=\prod_{k=1}^{d}\ell_{i_{k}}(x^{k})$.  
From the property $\ell_{i}(x_{j})=\delta_{ij}$ we have the corresponding multidimensional version $\ell_{\bi}(\bx_{\bj})=\delta_{\bi\bj}$, so that $\bU_{\DG}(\bx_{\bi},t)=\bU_{\bi}(t)$; i.e., the expansion coefficients in Equation \eqref{eq:conservedNodalExpansion} represent the conserved variables defined in the nodes $\bx_{\bi}$.  
To evaluate the integrals in Equation~\eqref{eq:semidiscreteDG}, we introduce numerical quadratures.  
First we define the one-dimensional $N$-point quadrature $Q_{N}^{i}:C^{0}(I^{i})\to\bbR$ with abscissas $\{\eta_{q}\}_{q=1}^{N}$ and weights $\{w_{q}\}_{q=1}^{N}$, normalized such that $\sum_{q=1}^{N}w_{q}=1$.  
(For example, the $N$-point Legendre-Gauss quadrature, which we will use, integrates polynomials of degree $\le 2N-1$ exactly.)
If $P(x)$ is such a polynomial, we have
\begin{equation}
  \f{1}{\Delta x}\int_{K}P(x)\,dx=\int_{I}P(\eta)\,d\eta=Q_{N}\big[P\big]\equiv\sum_{q=1}^{N}w_{q}\,P(\eta_{q}).  
\end{equation}

@@

\section{Euler Equations in Commonly Used Curvilinear Coordinates}
\label{app:CurvilinearEuler}

@@

\begin{equation}
  \pderiv{\rho}{t}
  +\f{1}{R}\pderiv{}{R}\Big(R\,\rho\,v_{R}\Big)
  +\pderiv{}{z}\Big(\rho\,v_{z}\Big)
  +\f{1}{R}\pderiv{}{\phi}\Big(\rho\,v_{\phi}\Big)=0,
\end{equation}
\begin{equation}
  \pderiv{\big(\rho\,v_{R}\big)}{t}
  +\f{1}{R}\pderiv{}{R}\Big(R\,\big(\rho\,v_{R}\,v_{R}+p\big)\Big)
  +\pderiv{}{z}\Big(\rho\,v_{R}\,v_{z}\Big)
  +\f{1}{R}\pderiv{}{\phi}\Big(\rho\,v_{R}\,v_{\phi}\Big)
  =\f{\big(\rho\,v_{\phi}^{2}+p\big)}{R},
\end{equation}
\begin{equation}
  \pderiv{\big(\rho\,v_{z}\big)}{t}
  +\f{1}{R}\pderiv{}{R}\Big(R\,\rho\,v_{z}\,v_{R}\Big)
  +\pderiv{}{z}\Big(\big(\rho\,v_{z}\,v_{z}+p\big)\Big)
  +\f{1}{R}\pderiv{}{\phi}\Big(\rho\,v_{z}\,v_{\phi}\Big)=0,
\end{equation}
\begin{equation}
  \pderiv{\big(\rho\,v_{\phi}\big)}{t}
  +\f{1}{R}\pderiv{}{R}\Big(R\,\rho\,v_{\phi}\,v_{R}\Big)
  +\pderiv{}{z}\Big(\rho\,v_{\phi}\,v_{z}\Big)
  +\f{1}{R}\pderiv{}{\phi}\Big(\rho\,v_{\phi}\,v_{\phi}+p\Big)
  =-\f{\rho\,v_{\phi}\,v_{R}}{R},
\end{equation}
\begin{equation}
  \pderiv{E}{t}+\f{1}{R}\pderiv{}{R}\Big(\,R\,\big(E+p\big)\,v_{R}\,\Big)+\pderiv{}{z}\Big(\,\big(E+p\big)\,v_{z}\,\Big)+\f{1}{R}\pderiv{}{\phi}\Big(\,\big(E+p\big)\,v_{\phi}\,\Big)=0.
\end{equation}
\begin{equation}
  \pderiv{n_{e}}{t}
  +\f{1}{R}\pderiv{}{R}\Big(R\,n_{e}\,v_{R}\Big)
  +\pderiv{}{z}\Big(n_{e}\,v_{z}\Big)
  +\f{1}{R}\pderiv{}{\phi}\Big(n_{e}\,v_{\phi}\Big)=0,
\end{equation}

>>>>

\newcommand{\ee}[1]{{\color{red} #1}}
\newcommand{\dx}{\Delta x}
\newcommand{\pbK}{\partial\mathbf{K}}
\newcommand{\sumx}{\sum_{i=1}^{d}}

@@

The model given by Eqs.~\eqref{eq:massConservation} - \eqref{eq:electronConservation} is non-relativistic (i.e., it does not include effects due to strong gravitational fields or relativistic fluid velocities), but is expressed in terms of curvilinear coordinates through the spatial metric tensor $\gamma_{ij}$, which gives squared infinitesimal line-element
\begin{equation}
  ds_{\vect{x}}^{2}=\gamma_{ij}\,dx^{i}\,dx^{j},
\end{equation}
where $\gamma_{ij}$ are the spatial components of the coordinate basis metric tensor.  
We will make two simplifying assumptions: the metric tensor is (1) diagonal and (2) time-independent.  
Specifically we assume the following diagonal form
\begin{equation}
  \gamma_{ij}
  =\mbox{diag}\big[\,\gamma_{11},\,\gamma_{22}(x^{1}),\,\gamma_{33}(x^{1},x^{2})\,\big], 
  \label{eq:threeMetric}
\end{equation}
which is sufficiently general to accommodate Cartesian, cylindrical, and spherical polar coordinates (see Table~\ref{tab:metricFunctions} and Appendix~\ref{app:CurvilinearEuler}).  
Moreover, $\gamma$ is the determinant of the metric tensor; $\sqrt{\gamma}=\sqrt{\gamma_{11}\gamma_{22}\gamma_{33}}$.  
The inverse of the spatial metric is denoted $\gamma^{ij}$, so that $\gamma^{ik}\gamma_{kj}=\delta^{i}_{~j}$.  
We can use the spatial metric to raise and lower indices on spatial vectors and tensors; e.g., $S_{j}=\gamma_{jk}\,S^{k}$.

\begin{table}
  \begin{center}
  \caption{Relevant metric functions for Cartesian, Cylindrical, and Spherical coordinate systems.\label{tab:metricFunctions}}
  \begin{tabular}{ccccccccccc}
    \midrule
    Coordinates & $x^{1}$ & $x^{2}$ & $x^{3}$ & $\gamma_{11}$ & $\gamma_{22}$ & $\gamma_{33}$ & $\sqrt{\gamma}$
    & $\f{1}{\gamma_{22}}\pderiv{\gamma_{22}}{x^{1}}$ & $\f{1}{\gamma_{33}}\pderiv{\gamma_{33}}{x^{1}}$ & $\f{1}{\gamma_{33}}\pderiv{\gamma_{33}}{x^{2}}$ \\
    \midrule
    \midrule
    Cartesian  & $x$ & $y$ & $z$ & 1 & 1 & 1 & 1 & 0 & 0 & 0 \\
    Cylindrical & $R$ & $z$ & $\phi$ & 1 & 1 & $R^{2}$ & $R$ & 0 & $2/R$ & 0 \\
    Spherical   & $r$ & $\theta$ & $\phi$ & 1 & $r^{2}$ & $r^{2}\sin^{2}\theta$ & $r^{2}\sin\theta$ & $2/r$ & $2/r$ & $2\cot\theta$ \\
    \midrule
    \midrule
  \end{tabular}
  \end{center}
\end{table}

@@

We define the geometry sources in Eq.~\eqref{eq:momentumConservation} by
\begin{equation}
  G_{j}
  =\f{1}{2}\,P^{ik}\,\pderiv{\gamma_{ik}}{x^{j}}-\rho\,\pderiv{\Phi}{x^{j}}.  
\end{equation}
Specifically, using the time-independent metric tensor in Eq.~\eqref{eq:threeMetric} and writing in terms of $P^{i}_{~j}$, the components are
\begin{equation}
  G_{1}
  =\f{1}{2}\,\Big(\,P^{2}_{~2}\,\f{1}{\gamma_{22}}\pderiv{\gamma_{22}}{x^{1}}+P^{3}_{~3}\,\f{1}{\gamma_{33}}\pderiv{\gamma_{33}}{x^{1}}\,\Big)-\rho\,\pderiv{\Phi}{x^{1}}, \quad
  G_{2}
  =\f{1}{2}\,P^{3}_{~3}\,\f{1}{\gamma_{33}}\pderiv{\gamma_{33}}{x^{2}}-\rho\,\pderiv{\Phi}{x^{2}}, \quad\text{and}\quad
  G_{3}
  =-\rho\,\pderiv{\Phi}{x^{3}}.
\end{equation}
The gravitational energy source term in Eq.~\eqref{eq:energyConservation} is denoted with $G_{E}=-S^{i}\,\pd{\Phi}{x^{i}}$.

Then, by defining the vector of evolved variables $\bU=\big(\,\rho,\,S_{1},\,S_{2},\,S_{3},\,E,\,n_{e}\,\big)^{T}$, the vector of geometry sources $\bG(\bU)=\big(\,0,\,G_{1},\,G_{2},\,G_{3},\,G_{E},\,0\,\big)^{T}$, and the flux vector components
\begin{equation}
  \vect{F}^{i}(\bU)
  =\big(\,\rho\,v^{i},\,P_{~1}^{i},\,P_{~2}^{i},\,P_{~3}^{i},\,(E+p)\,v^{i},\,n_{e}\,v^{i}\,\big)^{T},
\end{equation}
we write the system of equations in the usual compact form
\begin{equation}
  \pd{\bU}{t}
  +\f{1}{\sqrt{\gamma}}\pderiv{}{x^{i}}\Big(\,\sqrt{\gamma}\,\vect{F}^{i}(\bU)\,\Big)
  =\bG(\bU).
  \label{eq:extendedEulerCompact}
\end{equation}
Eq.~\eqref{eq:extendedEulerCompact} forms the basis for developing DG methods for the Euler equations.

@@

In the DG method, the conserved quantities are approximated by a local expansion of the form
\begin{equation}
  \bU_{\DG}(\bx,t)
  =\sum_{\bi=\vect{1}}^{\bN}\bU_{\bi}(t)\,\ell_{\bi}(\bx), 
\end{equation}
where basis functions $\ell_{\bi}(\bx)$, which we will take to be polynomials, belong to a function space denoted $\bbV^{k}$ and have local support in a computational cell or element denoted by $\bK$.

Several books and review articles on the DG method are now available \citep[see, e.g.,][]{cockburnShu_2001,hesthavenWarburton_2008}, and we will not go into too much detail here.  
However, we review some key concepts to introduce notation, and emphasize specific choices in our implementation.

@@

We divide the computational domain $D$ into a disjoint union $\mathcal{T}$ of open elements $\bK$, so that $D = \cup_{\bK \in \cT}\bK$.  
We require that each element is a box in the logical coordinates; i.e.,
\begin{equation}
  \bK=\{\vect{x} : x^{i} \in K^{i} := (\xL^{i},\xH^{i})\}, 
\end{equation}
with the surface elements denoted $\pbK^{i}=\otimes_{j\ne i}K^{j}$.  
We use $V_{\bK}$ to denote the proper volume of the element
\begin{equation}
  V_{\bK} = \int_{\bK}dV, \quad\text{where}\quad dV = \sqrt{\gamma}\,\prod_{i=1}^{d}dx^{i}.  
\end{equation}
We also define $\bx=\{\tilde{\bx}^{i},x^{i}\}$ and $\dx^{i}=\xH^{i}-\xL^{i}$.

We let the approximation space for the DG method, $\mathbb{V}^{k}$, to be constructed from the tensor product of one-dimensional polynomials of maximal degree $k$.  
Note that functions in $\mathbb{V}^{k}$ can be discontinuous across element interfaces.  
The semi-discrete DG problem is to find $\bU_{\DG}\in\mathbb{V}^{k}$ (which approximates $\bU$) such that
\begin{align}
  &\partial_{t}\int_{\bK}\bU_{\DG}\,v\,dV
  +\sumx\int_{\pbK^{i}}\big(\,\sqrt{\gamma}\,\widehat{\vect{F}}^{i}(\bU_{\DG})\,v\big|_{\xH^{i}}-\sqrt{\gamma}\,\widehat{\vect{F}}^{i}(\bU_{\DG})\,v\big|_{\xL^{i}}\,\big)\,\tilde{\bx}^{i} \nonumber \\
  &\hspace{24pt}
  -\sumx\int_{\bK}\vect{F}^{i}(\bU_{\DG})\,\pderiv{v}{x^{i}}\,dV
  =\int_{\bK}\bG(\bU_{\DG})\,v\,dV
  \label{eq:semidiscreteDG}
\end{align}
for all $v\in\mathbb{V}^{k}$ and all $\bK\in\mathcal{T}$.

To connect the elements in Eq.~\eqref{eq:semidiscreteDG}, $\widehat{\bF}^{i}(\bU_{\DG})$ is a numerical flux approximating the flux on the $i$th surface of $\bK$.  
For this purpose we define the numerical flux function $\vect{f}^{i}$, which evaluates the numerical flux given values from both sides of an element interface; i.e.,
\begin{equation}
  \widehat{\bF}^{i}(\bU_{\DG})=\vect{f}^{i}(\bU_{\DG}(x^{i,-},\tilde{\bx}^{i}),\bU_{\DG}(x^{i,+},\tilde{\bx}^{i})),
\end{equation}
where superscripts $-/+$, e.g., in the arguments of $\bU_{\DG}(x^{i,-/+},\tilde{\bx}^{i})$, indicate that the function is evaluated to the immediate left/right of $x^{i}$.  
For example, the simple Lax-Friedrichs flux is given by
\begin{equation}
  \vect{f}^{i}(\bU^{-},\bU^{+})
  =\f{1}{2}\,\big(\,\bF^{i}(\bU^{-})+\bF^{i}(\bU^{+})-\alpha^{i}\,(\,\bU^{+}-\bU^{-}\,)\,\big),
\end{equation}
where $\alpha^{i}=||\mbox{eig}\big(\partial\bF^{i}/\partial\bU\big)||_{\infty}$ is the largest eigenvalue of the flux jacobian.

The approximation space $\mathbb{V}^{k}$ contains the constant functions, and the choice $v=1$ in Eq.~\eqref{eq:semidiscreteDG} gives
\begin{equation}
  \partial_{t}\bU_{\bK}
  +\f{1}{V_{\bK}}\sumx\int_{\pbK^{i}}\big(\,\sqrt{\gamma}\,\widehat{\vect{F}}^{i}(\bU_{\DG})\big|_{\xH^{i}}-\sqrt{\gamma}\,\widehat{\vect{F}}^{i}(\bU_{\DG})\big|_{\xL^{i}}\,\big)\,d\tilde{\bx}^{i}
  =\bG_{\bK},
\end{equation}
where we have defined the volume averages
\begin{equation}
  \bU_{\bK}=\f{1}{V_{\bK}}\int_{\bK}\bU_{\DG}\,dV \quad\text{and}\quad
  \bG_{\bK}=\f{1}{V_{\bK}}\int_{\bK}\bG(\bU_{\DG})\,dV.  
\end{equation}
This illustrates how the cell average (similar to finite volume methods) is evolved in the DG method.

@@

Here we provide further details on the DG method in order to arrive at the equations that are actually implemented in our code.  
We start by introducing some notation, defining the polynomial expansion for $\bcM_{\DG}$, and the quadrature rules used to evaluate the integrals in Equation~\eqref{eq:semidiscreteDG}.  
Then, using these rules, we provide explicit expressions for each of the terms in Equation~\eqref{eq:semidiscreteDG}.

In each element $\bK$, we use a nodal representation in the conserved variables $\bU$; i.e.,
\begin{equation}
  \bU(\bx,t)\approx
  \bU_{\DG}(\bx,t)=\sum_{\bi=\vect{1}}^{\bN}\bU_{\bi}(t)\,\ell_{\bi}(\bx),
  \label{eq:conservedNodalExpansion}
\end{equation}
where $\ell_{\bi}(\bx)\in\bbQ^{k}$ are basis functions.  
Specifically, we use one-dimensional Lagrange polynomials $\ell_{i}(x)$ to construct the multidimensional representation, where
\begin{equation}
  \ell_{i}(\eta)=
  \prod_{\substack{j=1\\j\ne i}}^{N}\f{\eta-\eta_{j}}{\eta_{i}-\eta_{j}}.  
\end{equation}
Moreover $\bi=\{i_{1},\ldots,i_{d}\}$ is a multi-index, and $\ell_{\bi}(\bx)=\prod_{k=1}^{d}\ell_{i_{k}}(x^{k})$.  
From the property $\ell_{i}(x_{j})=\delta_{ij}$ we have the corresponding multidimensional version $\ell_{\bi}(\bx_{\bj})=\delta_{\bi\bj}$, so that $\bU_{\DG}(\bx_{\bi},t)=\bU_{\bi}(t)$; i.e., the expansion coefficients in Equation \eqref{eq:conservedNodalExpansion} represent the conserved variables defined in the nodes $\bx_{\bi}$.  
To evaluate the integrals in Equation~\eqref{eq:semidiscreteDG}, we introduce numerical quadratures.  
First we define the one-dimensional $N$-point quadrature $Q_{N}^{i}:C^{0}(I^{i})\to\bbR$ with abscissas $\{\eta_{q}\}_{q=1}^{N}$ and weights $\{w_{q}\}_{q=1}^{N}$, normalized such that $\sum_{q=1}^{N}w_{q}=1$.  
(For example, the $N$-point Legendre-Gauss quadrature, which we will use, integrates polynomials of degree $\le 2N-1$ exactly.)
If $P(x)$ is such a polynomial, we have
\begin{equation}
  \f{1}{\Delta x}\int_{K}P(x)\,dx=\int_{I}P(\eta)\,d\eta=Q_{N}\big[P\big]\equiv\sum_{q=1}^{N}w_{q}\,P(\eta_{q}).  
\end{equation}

@@

\section{Euler Equations in Commonly Used Coordinate Systems}
\label{app:CurvilinearEuler}

@@

Mass conservation equation
\begin{equation}
  \pderiv{\rho}{t}
  +\f{1}{R}\pderiv{}{R}\Big(R\,\rho\,v^{1}\Big)
  +\pderiv{}{z}\Big(\rho\,v^{2}\Big)
  +\pderiv{}{\phi}\Big(\rho\,v^{3}\Big)=0,
\end{equation}
momentum conservation equation ($R$-component)
\begin{equation}
  \pderiv{\big(\rho\,v_{1}\big)}{t}
  +\f{1}{R}\pderiv{}{R}\Big(R\,\big(\rho\,v_{1}\,v^{1}+p\big)\Big)
  +\pderiv{}{z}\Big(\rho\,v_{1}\,v^{2}\Big)
  +\pderiv{}{\phi}\Big(\rho\,v_{1}\,v^{3}\Big)
  =\f{\big(\rho\,v_{3}\,v^{3}+p\big)}{R}-\rho\pderiv{\Phi}{R},
\end{equation}
momentum conservation equation ($z$-component)
\begin{equation}
  \pderiv{\big(\rho\,v_{2}\big)}{t}
  +\f{1}{R}\pderiv{}{R}\Big(R\,\rho\,v_{2}\,v^{1}\Big)
  +\pderiv{}{z}\Big(\rho\,v_{2}\,v^{2}+p\Big)
  +\pderiv{}{\phi}\Big(\rho\,v_{2}\,v^{3}\Big)=-\rho\pderiv{\Phi}{z},
\end{equation}
momentum conservation equation ($\phi$-component)
\begin{equation}
  \pderiv{\big(\rho\,v_{3}\big)}{t}
  +\f{1}{R}\pderiv{}{R}\Big(R\,\rho\,v_{3}\,v^{1}\Big)
  +\pderiv{}{z}\Big(\rho\,v_{3}\,v^{2}\Big)
  +\pderiv{}{\phi}\Big(\rho\,v_{3}\,v^{3}+p\Big)
  =-\rho\pderiv{\Phi}{\phi},
\end{equation}
energy equation
\begin{equation}
  \pderiv{E}{t}+\f{1}{R}\pderiv{}{R}\Big(\,R\,\big(E+p\big)\,v^{1}\,\Big)+\pderiv{}{z}\Big(\,\big(E+p\big)\,v^{2}\,\Big)+\pderiv{}{\phi}\Big(\,\big(E+p\big)\,v^{3}\,\Big)=-\rho\,v^{i}\pderiv{\Phi}{x^{i}},
\end{equation}
electron conservation equation
\begin{equation}
  \pderiv{n_{e}}{t}
  +\f{1}{R}\pderiv{}{R}\Big(R\,n_{e}\,v^{1}\Big)
  +\pderiv{}{z}\Big(n_{e}\,v^{2}\Big)
  +\pderiv{}{\phi}\Big(n_{e}\,v^{3}\Big)=0.  
\end{equation}
